---
title: "Comparison of segment frequencies in BDPROTO, PHOIBLE and SegBo"
author: "Steven Moran\n"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

# Overview

Plot the raw frequency distributions from the three databases.

Load some R libraries.

```{r}
library(tidyverse)
library(knitr)
```

Load the aggregated and transformed data created by the [data preparation report](../data_preparation/data_prep.md).

```{r}
load('../data_preparation/dfs_for_analysis_csv_vs.RData')
```

The `all_segments_long` table looks like this:

```{r}
all_segments_long %>% head() %>% kable()
```

The table is already ordered by PHOIBLE segment frequency, but R needs it sorted as factor to plot it sorted.

```{r}
all_segments_long$Phoneme <- factor(all_segments_long$Phoneme, levels=all_segments_long$Phoneme[order(-all_segments_long$phoible)])
```

# Plots

Let's plot the top top 30 most frequent phonemes in PHOIBLE.

```{r}
top.z <- head(all_segments_long, n=30)

ggplot(data = top.z, aes(x = Phoneme, group=1)) +
  geom_line(aes(y = phoible_freq, color = "PHOIBLE")) + 
  geom_line(aes(y = bdproto_freq, color = "BDPROTO")) +
  geom_line(aes(y = segbo_freq, color = "SegBo")) +
  ylab("Frequency of phoneme in database") +
  # xlab("Thirty most frequent phonemes as reported in PHOIBLE") +
  labs(color="Database") +
  theme_bw()
```

Get the most frequent phonemes in BDPROTO and plot them against PHOIBLE.

```{r}
bdproto_top <- all_segments_long %>% arrange(desc(bdproto))
```

Ordered as factors for plotting.

```{r}
bdproto_top$Phoneme <- factor(bdproto_top$Phoneme, levels=bdproto_top$Phoneme[order(-bdproto_top$bdproto)])
```

```{r}
top.z <- head(bdproto_top, n=30)

ggplot(data = top.z, aes(x = Phoneme, group=1)) +
  geom_line(aes(y = phoible_freq, color = "PHOIBLE")) + 
  geom_line(aes(y = bdproto_freq, color = "BDPROTO")) +
  geom_line(aes(y = segbo_freq, color = "SegBo")) +
  ylab("Frequency of phoneme in database") +
  # xlab("Thirty most frequent phonemes as reported in BDPROTO") +
  labs(color="Database") +
  theme_bw()
  # theme(legend.position = "none") 
```

Get the most frequent phonemes in SegBo and plot them against PHOIBLE and BDPROTO.

```{r}
segbo_top <- all_segments_long %>% arrange(desc(segbo))
```

Ordered as factors for plotting.

```{r}
segbo_top$Phoneme <- factor(segbo_top$Phoneme, levels=segbo_top$Phoneme[order(-segbo_top$segbo)])
```

```{r}
top.z <- head(segbo_top, n=30)

ggplot(data = top.z, aes(x = Phoneme, group=1)) +
  geom_line(aes(y = segbo_freq, color = "SegBo")) +
  geom_line(aes(y = phoible_freq, color = "PHOIBLE")) + 
  geom_line(aes(y = bdproto_freq, color = "BDPROTO")) +

  ylab("Frequency of phoneme in database") +
  # xlab("Thirty most frequent phonemes as reported in BDPROTO") +
  labs(color="Database") +
  theme_bw()
  # theme(legend.position = "none") 
```

# Tables

Let's generate a summary table for the paper that includes the number of languages, language families, and total segments in each database.

```{r}
# Get number of phonological inventories
phoible_ids <- all_dbs_all_segments %>% filter(Database=="phoible") %>% select(InventoryID) %>% group_by(InventoryID) %>% summarize(n=n())
phoible_num_ids <- nrow(phoible_ids)

bdproto_ids <- all_dbs_all_segments %>% filter(Database=="bdproto") %>% select(InventoryID) %>% group_by(InventoryID) %>% summarize(n=n())
bdproto_num_ids <- nrow(bdproto_ids)

segbo_ids <- all_dbs_all_segments %>% filter(Database=="segbo") %>% select(InventoryID) %>% group_by(InventoryID) %>% summarize(n=n())
segbo_num_ids <- nrow(segbo_ids)

# Generate segment counts
phoible_segments <- all_dbs_all_segments %>% filter(Database=="phoible") %>% select(Phoneme) %>% group_by(Phoneme) %>% summarize(n=n())
phoible_num_segments <- nrow(phoible_segments)

bdproto_segments <- all_dbs_all_segments %>% filter(Database=="bdproto") %>% select(Phoneme) %>% group_by(Phoneme) %>% summarize(n=n())
bdproto_num_segments <- nrow(bdproto_segments)

segbo_segments <- all_dbs_all_segments %>% filter(Database=="segbo") %>% select(Phoneme) %>% group_by(Phoneme) %>% summarize(n=n())
segbo_num_segments <- nrow(segbo_segments)

# Generate language family counts based on unique Glottocodes
phoible_langs <- all_dbs_all_segments %>% filter(Database=="phoible") %>% select(Glottocode) %>% group_by(Glottocode) %>% summarize(n=n())
phoible_num_langs <- nrow(phoible_langs)

bdproto_langs <- all_dbs_all_segments %>% filter(Database=="bdproto") %>% select(Glottocode) %>% group_by(Glottocode) %>% summarize(n=n())
bdproto_num_langs <- nrow(bdproto_langs)

segbo_langs <- all_dbs_all_segments %>% filter(Database=="segbo") %>% select(Glottocode) %>% group_by(Glottocode) %>% summarize(n=n())
segbo_num_langs <- nrow(segbo_langs)

# Generate language counts based on unique Glottocodes (language varieties)
phoible_lang_fams <- all_dbs_all_segments %>% filter(Database=="phoible") %>% select(family_id) %>% group_by(family_id) %>% summarize(n=n())
phoible_num_lang_fams <- nrow(phoible_lang_fams)

bdproto_lang_fams <- all_dbs_all_segments %>% filter(Database=="bdproto") %>% select(family_id) %>% group_by(family_id) %>% summarize(n=n())
bdproto_num_lang_fams <- nrow(bdproto_lang_fams)

segbo_lang_fams <- all_dbs_all_segments %>% filter(Database=="segbo") %>% select(family_id) %>% group_by(family_id) %>% summarize(n=n())
segbo_num_lang_fams <- nrow(segbo_lang_fams)

```

Create the table.

```{r}
Database <- c("BDPROTO", "PHOIBLE", "SegBo")
inventories <- c(bdproto_num_ids, phoible_num_ids, segbo_num_ids)
langs <- c(bdproto_num_langs, phoible_num_langs, segbo_num_langs)
fams <- c(bdproto_num_lang_fams, phoible_num_lang_fams, segbo_num_lang_fams)
segments <- c(bdproto_num_segments, phoible_num_segments, segbo_num_segments)
t <- data.frame(Database, inventories, langs, fams, segments)
colnames(t) <- c('Database', 'Phonological inventories', 'Language varieties', 'Language families', 'Segment types')
t %>% kable()
```

Let's dump it to LaTeX for the paper.

```{r}
library(xtable)
print(xtable(t), include.rownames=FALSE)
```